---
title: "Bayesian methods and modern statistics"
subtitle: "Fall 2023"
mainfont: Lato
format: 
   html: 
    page-layout: full
    toc: false
  
---

# Schedule

```{r load-libraries, echo = F, message = F, eval = F}
library(tidyverse)
library(knitr)
library(emo)
```

```{r smart-date-setup, echo = F, message = F, eval = F}
# set these:
firstDay = as.Date("2023-08-28")
firstDayID = "Mon"
lastDay = as.Date("2023-12-08")
lectureDays = c("Tue", "Thu")
labDays = c("Mon")
holidays = as.Date(c("2023-09-04", "2023-10-16", "2023-10-17", "2023-11-22", "2023-11-23", "2023-11-24"))

### render table
allDates = seq(from = firstDay, to = lastDay, by = 'day')
schedule = data.frame(allDates,
                        dayOfWeek = weekdays(allDates, abbreviate=T)) %>%
  filter(dayOfWeek %in% c(lectureDays, labDays))
schedule = schedule %>%
  mutate(lab = ifelse(dayOfWeek %in% labDays, TRUE, FALSE)) %>%
  mutate(Topic = ifelse(allDates %in% holidays, "NO CLASS", "")) %>%
  mutate(Date = format(allDates, format = "%b %d")) %>%
  mutate(Date = paste(dayOfWeek, Date)) %>%
  mutate(Week = rep("", nrow(schedule))) %>%
  mutate(Prepare = rep("", nrow(schedule))) %>%
  mutate(AE = rep("", nrow(schedule))) %>%
  mutate(Assignment = rep("", nrow(schedule))) %>%
  mutate(Exam = rep("", nrow(schedule))) %>%
  mutate(Project = rep("", nrow(schedule))) %>%
  select(Week, Date, Topic, Prepare, AE, Assignment, Exam, Project)
# to be pasted into console then pasted into markdown:
schedule %>%
  kable(format = "markdown")
```

| Week | Date       | Topic                        | Reading | Notes | Assignment |
|:-----|:-----------|:----------------------------------|-----|:------:|:-----------|
| 1    | Mon Aug 28 | welcome                     |         | [💻](/slides/lab0-welcome.html)       | [hello R](/labs/lab0.html)           |
|      | Tue Aug 29 | intro, history, notation          | Ch. 2   |       | [hw 0](/hw/hw00.html)            |
|      | Thu Aug 31 | probability, exchangeability | Ch. 2   |[💻](/notes/probability.html)       |  [hw 1](/hw/hw01.html)          |
| 2    | Mon Sep 04 | NO CLASS                     |         |       |            |
|      | Tue Sep 05 | single parameter estimation                             | Ch. 3        | [💻](/notes/estimation1.html) |       |            |
|      | Thu Sep 07 | Poisson model and conjugacy                             | Ch. 3         |        | [hw 2](/hw/hw02.html)           |
| 3    | Mon Sep 11 | MLE and MAP estimator                              |         | [💻](/slides/lab1.html)      |            |
|      | Tue Sep 12 | Reliability, exp. families, prediction                            | Ch. 3        | [💻](/notes/prediction1.html)      |            |
|      | Thu Sep 14 |  Monte Carlo integration                         | Ch. 4        |       |            |
| 4    | Mon Sep 18 |                              |         |       |            |
|      | Tue Sep 19 | the normal model                            | Ch. 5        |       |            |
|      | Thu Sep 21 | two parameter estimation                           | Ch. 5        |       |            |
| 5    | Mon Sep 25 |                              |         |       |            |
|      | Tue Sep 26 | catch up / review                            |         |       |            |
|      | Thu Sep 28 | Exam I                             |         |       |            |
| 6    | Mon Oct 02 |                              |         |       |            |
|      | Tue Oct 03 | Gibbs sampling                             |         |       |            |
|      | Thu Oct 05 |                              |         |       |            |
| 7    | Mon Oct 09 |                              |         |       |            |
|      | Tue Oct 10 |                              |         |       |            |
|      | Thu Oct 12 |                              |         |       |            |
| 8    | Mon Oct 16 | NO CLASS                     |         |       |            |
|      | Tue Oct 17 | NO CLASS                     |         |       |            |
|      | Thu Oct 19 |                              |         |       |            |
| 9    | Mon Oct 23 |                              |         |       |            |
|      | Tue Oct 24 |                              |         |       |            |
|      | Thu Oct 26 |                              |         |       |            |
| 10   | Mon Oct 30 |                              |         |       |            |
|      | Tue Oct 31 |                              |         |       |            |
|      | Thu Nov 02 |                              |         |       |            |
| 11   | Mon Nov 06 |                              |         |       |            |
|      | Tue Nov 07 |                              |         |       |            |
|      | Thu Nov 09 |                              |         |       |            |
| 12   | Mon Nov 13 |                              |         |       |            |
|      | Tue Nov 14 |                              |         |       |            |
|      | Thu Nov 16 |  Exam II                            |         |       |            |
| 13   | Mon Nov 20 |                              |         |       |            |
|      | Tue Nov 21 |                              |         |       |            |
|      | Thu Nov 23 | NO CLASS                     |         |       |            |
| 14   | Mon Nov 27 |                              |         |       |            |
|      | Tue Nov 28 |                              |         |       |            |
|      | Thu Nov 30 |                              |         |       |            |
| 15   | Mon Dec 04 |                              |         |       |            |
|      | Tue Dec 05 |                              |         |       |            |
|      | Thu Dec 07 |                              |         |       |            |

<!-- | Week | Date       | Topic                                  |              Notes              | Lab                      | Exam                                  | Project | -->

<!-- |:----------|:----------|:----------|:---------:|:----------|:----------|:----------| -->

<!-- | 0    | Fri Jan 13 | welcome                                |  [💻](/slides/01-welcome.html)  |                          |                                       |         | -->

<!-- | 1    | Tue Jan 17 | git                                    | [💻](/slides/lab0-welcome.html) | [lab 0](/labs/lab0.html) |                                       |         | -->

<!-- |      | Wed Jan 18 | control flow, loops and functions      |    [💻](/slides/lec02.html)     |                          |                                       |         | -->

<!-- |      | Fri Jan 20 | NAs, lists, NULL and attributes        |    [💻](/slides/lec03.html)     |                          |                                       |         | -->

<!-- | 2    | Tue Jan 24 | if it's a function, fix it             |                                 | [lab 1](/labs/lab1.html) |                                       |         | -->

<!-- |      | Wed Jan 25 | object oriented programming            |    [💻](/slides/lec04.html)     |                          |                                       |         | -->

<!-- |      | Fri Jan 27 | tidy data and dplyr                    |    [💻](/slides/lec05.html)     |                          |                                       |         | -->

<!-- | 3    | Tue Jan 31 | computational masonry                  |                                 | [lab 2](/labs/lab2.html) |                                       |         | -->

<!-- |      | Wed Feb 01 | subsetting                             |    [💻](/slides/lec06.html)     |                          |                                       |         | -->

<!-- |      | Fri Feb 03 | ggplot2                                |    [💻](/slides/lec07.html)     |                          |                                       |         | -->

<!-- | 4    | Tue Feb 07 | dream of plots                         |                                 | [lab 3](/labs/lab3.html) |                                       |         | -->

<!-- |      | Wed Feb 08 | functional programming and purrr       |    [💻](/slides/lec08.html)     |                          |                                       |         | -->

<!-- |      | Fri Feb 10 | regular expressions                    |    [💻](/slides/lec09.html)     |                          |                                       |         | -->

<!-- | 5    | Tue Feb 14 | wrangling the written word             |                                 | [lab 4](/labs/lab4.html) |                                       |         | -->

<!-- |      | Wed Feb 15 | catch-up                               |                                 |                          |                                       |         | -->

<!-- |      | Fri Feb 17 | review / questions / exam              |                                 |                          | [exam 1 release](/exams/exam01.html) |         | -->

<!-- | 6    | Tue Feb 21 | no lab: exam 1                         |                                 |                |   exam 1 due                                    |         | -->

<!-- |      | Wed Feb 22 | complexity and parallelization |  [💻](/slides/lec10.html)                                |                          |                                       |         | -->

<!-- |      | Fri Feb 24 | intro to optimization          |  [💻](/slides/lec11.html)                                |                          |                                       |         | -->

<!-- | 7    | Tue Feb 28 | where's zero?                                       |                                 | [lab 5](/labs/lab5.html)                         |                                       |         | -->

<!-- |      | Wed Mar 01 | Newton-Raphson                  |[💻](/slides/lec12.html)                                  |                          |                                       |         | -->

<!-- |      | Fri Mar 03 | Newton-Raphson examples           |                                 |                          |                                       |         | -->

<!-- | 8    | Tue Mar 07 |  continue lab                                      |                                 |                          |                                       |         | -->

<!-- |      | Wed Mar 08 | MM principles   |   [💻](/slides/lec13.html)                              |                          |                                       |         | -->

<!-- |      | Fri Mar 10 | MM example                        |   [💻](/slides/lec14.html)                              |                          |                                       |         | -->

<!-- | 9    | Tue Mar 14 | NO CLASS: Spring break                 |                                 |                          |                                       |         | -->

<!-- |      | Wed Mar 15 | NO CLASS: Spring break                 |                                 |                          |                                       |         | -->

<!-- |      | Fri Mar 17 | NO CLASS: Spring break                 |                                 |                          |                                       |         | -->

<!-- | 10   | Tue Mar 21 | ascend the likelihood                                      |                                 | [lab 6](/labs/lab6.html)                          |                                       |         | -->

<!-- |      | Wed Mar 22 | EM algorithm                      | [💻](/slides/lec15.html)                                 |                          |                                       |         | -->

<!-- |      | Fri Mar 24 | Monte Carlo integration                          |  [💻](/slides/lec16.html)                                |                          |                                       |         | -->

<!-- | 11   | Tue Mar 28 | continue lab                                       |                                 |                          |                                       |         | -->

<!-- |      | Wed Mar 29 | importance sampling                                 | [💻](/slides/lec17.html)                                 |                          |                                       |         | -->

<!-- |      | Fri Mar 31 | review / questions / exam                          |                                 |                          |   [exam 2 release](/exams/exam02.html)                                     |          | -->

<!-- | 12   | Tue Apr 04 |  no lab: exam 2                                       |                               |                          |  exam 2 due                                     |         | -->

<!-- |      | Wed Apr 05 |  shiny I                 |  [💻](/slides/lec18.html)                                |                          |                                       |         | -->

<!-- |      | Fri Apr 07 | shiny II              |                                 |                          |                                       |         | -->

<!-- | 13   | Tue Apr 11 | shiny dashboard                                  |                                 |               [on GitHub](https://github.com/orgs/sta323-sp23/repositories)         |                                  |         | -->

<!-- |      | Wed Apr 12 | SQL                 |                                 |                          |                                       |         | -->

<!-- |      | Fri Apr 14 | project tips                          |                                 |                          |                                       |         | -->

<!-- | 14   | Tue Apr 18 | project workday                                     |                                 |                          |                                       |         | -->

<!-- |      | Wed Apr 19 | APIs                                   |   [💻](/slides/lec20.html)                               |                          |                                       |         | -->

<!-- |      | Fri Apr 21 | covariance & PCA                                    | [💻](/slides/lec21.html)                                |                          |                                       |         | -->

<!-- | 15   | Tue Apr 25 | project presentations              |                                 |                          |                                       |         | -->

<!-- |      | Wed Apr 26 | special topic: cryptanalysis                         |  [💻](/slides/lec22.html)                               |                          |                                       |         | -->
